*Prochlorococcus marinus* responses to light and oxygen

### 2024-06-10 Output/PLOSRound2/ResponsetoReviewers.txt

2. Please note that PLOS ONE has specific guidelines on code sharing for submissions in which author-
generated code underpins the findings in the manuscript. In these cases, all author-generated code 
must be made available without restrictions upon publication of the work. Please review our guidelines 
at https://journals.plos.org/plosone/s/materials-and-software-sharing#loc-sharing-code and ensure that 
your code is shared in a way that follows best practice and facilitates reproducibility and reuse.

3. Thank you for uploading your study's underlying data set. Unfortunately, the repository you have 
noted in your Data Availability statement does not qualify as an acceptable data repository according to 
PLOS's standards.

At this time, please upload the minimal data set necessary to replicate your study's findings to a stable, 
public repository (such as figshare or Dryad) and provide us with the relevant URLs, DOIs, or accession 
numbers that may be used to access these data. For a list of recommended repositories and additional 
information on PLOS standards for data deposition, please see 
https://journals.plos.org/plosone/s/recommended-repositories.

All data and annotated code for data import, transformations and analyses are
available at are available on the figshare data repository (DOI: 10.6084/m9.figshare.25959061).

@@

Line 55: LLII/III and LLIV are not the ?oldest? clades, but the most basal or early branching.

-We edited the text in the revised manuscript: "Clades LLII and LLIII, including cultured strain SS120, are
grouped together as early branching phylogenetic lineage in
the P. marinus radiation, with a preference for low light."
Line: 64

Line 369: The authors invoke ?limitations in the annotation process? to explain the lack of evidence for 
MED4 ATP synthase. What does this mean? The MED4 genome surely contains ATP synthase genes.

-We agree that the MED4 genome includes genes encoding ATP Synthase. Upon verifying our annotation assignments we discovered the problem with the absence of ATP Synthase from Clade HLI was not due to the limitations of the process but rather a mislabel of the column in figure 2. 'ATP Synthase' should have been labelled 'RUBISCO'
and 'RUBISCO' should have been labelled 'ATP Synthase'. However, in the Ocean Protein Portal database no RUBISCO peptides are annotated as coming from MED4. This likely reflects an annotation assignment issue because the RUBISCO proteins are highly conserved, so assignments to strains or clades of Prochlorococcus may be ambiguous.
We edited figure #2 to add 'jitter' offsets up to 15% of full axes scales, to visualize over-laid data points. 
We edited the text in the revised manuscript to reflect the limitation of the annotation process as it pertains to the highly conserved RUBISCO complex: "Though present and expressed in the genome of MED4 [65], the absence of proteins annotated as RUBISCO complex for clade HLI, compared to annotated detections of RUBISCO across the other three clades, suggests limitations in the annotation process assigning highly conserved protein sequences to clades."
Lines: 394 to 398

@@

Paragraph starting L363: Why is there no discussion of the Rubisco data? Or especially the absence 
thereof at low O2 for ecotypes HL1 and LL1? Since Rubisco is one of the most abundant proteins in a 
photosynthesizing cell, does this indicate that these ecotypes are present but not actively fixing carbon 
under these conditions? Also, is it worth noting that LLIV is the only variety to be detected at all in the 
deepest low O2 site -- which would appear to be consistent with subsequent findings in this study?

- We revised the text from lines 398 to 401 to incorporate these helpful observations. The following was added: "Notwithstanding limitations on assignments of proteins to clades, detections of the abundant carbon fixation complex RUBISCO derived from clades HLI, LLI and LLII/III were notably absent from stations at low [O2], suggesting limited capacity for carbon fixation by clades HLI, LLI and LLII/LLIII under low [O2] habitats."

Figures 3 through 8: These figures are obviously the crux of the results in this paper, but as they are 
currently presented it is relatively difficult to think clearly about them. I feel like they could be combined 
into 2 or 3 full page figures which might make it easier to make the relevant comparisons; either by 
combining the two types of figure (growth rates and model fits) for each ecotype, or else all the growth 
rate curves together in one figure and all the model plots together in another.

- Thank you for this suggestion. We combined the growth rate figure with the GAM model for each strain to facilitate comparisons.

@@

We briefly motivated the inclusion of the red light growth trials, as mechanistically informative, at line 432:
"As mentioned, growth rate trials under red light, although not representative of *P. marinus* niches, are mechanistically informative [@murphyPhotoinactivationPhotosystemII2017] regarding photoinactivation of PSII."

-Through the Discussion, notably under the discussion of the GAM models and responses to growth rate to increasing PUR, we return to the responses to red light to show that some 'high light' limitations on growth are interpretable as photoinhibition of PSII, which is more severe under blue, compared to red, light, notably at lines 646-650.

I also wonder if the data from the model fits could be used to predict which of the three ecotypes would 
prevail (have the highest growth rate) across the parameter space depicted, resulting in another figure 
showing something like "zones of exclusion" for each ecotype, which would be interesting in terms of 
predictive power. One could easily imagine applying those "zones of exclusion" to a map of the actual 
ocean at various depths, or even comparing current with projected future conditions of spreading 
OMZs. This is merely a suggestion -- I don't expect the authors to necessarily do this, but I think if they 
chose to it would make the manuscript of broader interest.

-Thank you for the comment. We added a new figure (Fig 10) of potential clade occupancy in temperate and equatorial seasonal and depth niches, in future warmed oceans under full oxygen or oxygen minimum zones.

@@

84: Follett et al 2022 PNAS showed that temp is not the only factor influencing poleward issues - 
ecological interactions may further impact poleward shift during warming

Thank you for  bringing this to our attention. We added the following text to the revised manuscript: "Follett *et al*. , however, model interactions of heterotrophic  bacteria  that may influence latitudinal expansions of *P. marinus*"
Line: 93

199: Could you please state more about why you are confident in this assignment? What validation or 
metric leads you to this?

Thank you for noting that this needs further clarification. 
We edited the paragraph for clarity in the revised manuscript adding: "While the accuracy of strain specific protein annotations is limited due to the high conservation of the target protein complexes, the two step approach of peptide-to-spectrum matching using deep paired metagenomics, does assign proteins at the level of clades."
Lines: 215-218

@@

Thank you for this comment. In this work we used 1 to 3 replicates depending on the experimental conditions. Specific number of replicates (shown as small dots in Figures (3 - 5)).
We edited the text in the revised manuscript:"Consistent absence of growth of some strains under some levels of photoperiod, PAR, or [O2] meant we completed 268 growth rate factor treatment combinations with 1 to 3 replicates."
Lines: 266-268

431, and generally throughout the paper: Would the authors care to comment on any connection 
between the differences in blue vs red light growth and the underlying amount of energy input from 
these different wavelengths?

As outlined in responses to questions from Review 1, the implications of blue vs. red light wavebands indeed relate to the higher energy per photon in the blue light band.  In particular blue light is absorbed well by chlorophyll, but is also directly absorbed by the manganese cluster of PSII, provoking one major path of photoinactivation, as discussed in Murphy et al. 2017, and cited at lines 307, 434, 648, 663, 676, and 894 .  In contrast, red light is absorbed well by chlorophyll, driving photosynthesis, but does not provoke direct photoinactivation of the PSII manganese cluster, lowering the metabolic cost of maintaining the photosynthetic system, compared to equivalent rates of photosynthesis driven by blue light.

These issues are now discussed in relation to figures 6 and S4, with a given amount of absorbed red light PUR provoking more growth than the equivalent absorbed blue light PUR.  In particular:

@@

Fig 4 and related figures: It might be useful to plot the specific measured points within the countour plot

Thank you for your suggestion. We combined the growth rate figure with the GAM model for each strain. We believe this will aid the reader with interpreting both plots. These are now figures 3 to 5.

685: The lack of detected FtsH could simply be a factor of low overall spectral counts and detection 
limits in the dataset; please provide more context to help the reader understand whether this could be 
just a small number problem with respect to MIT9313 in low O2 samples; Fig. 10 seems to just show a 
binary detected/not detected value.

Thank you for your suggestion. We agree that detection limits vary for different target proteins, particularly for detection of single proteins such as FtsH isoforms.  Such issues are less problematic for summed detections of multiple proteins assigned to multi-protein complexes such as PSII. An analyses of mass spectrophotometric proteomic detection limits is beyond the scope of the paper. We, however, edited figures 2 and 7 to add 'jitter' offsets up to 15% of full axes scales, to visualize multiple previously over-laid data points.

### 2024-07-02 Code/Manuscript_O2.Rmd
FIxed gif. cap for fig10

#  This is Fig10.tiff
NichesMap_cap <- glue("**Potential Future Niches for *Prochlorococcus* clades in a warming ocean.**",
                      "The maps shows current latitudinal distribution limits for *Prochlorococcus* (between green dashed lines), along with tropical and temperate  latitudinal regions (yellow bands).",
"Light levels, photoperiods and spectral bands, potentially permissive for growth of *Prochlorococcus* clades are determined by the interactions of latitude, season and depth attenuation of light.",
"**A.** Potential clade occupancies under Full Oxygen Zones (250 µM O~2~) are shown for 10-50 m and 100 m depths, in temperate vs. tropical regions, under depth-resolved photoperiods, approximating Winter Solstice, Equinox and Summer Solstice. Clade HLI is excluded from temperate depths by short photoperiods at Winter Solstice. Clades LLII/III and LLIV are excluded from shallow depths by excessive light.",
"**B.** Oxygen Minimum Zones (25 µM O~2~) drastically alter these potential occupancy patterns by excluding clade HLI from the temperate winter solstice at all depths, and by expanding the exploitation of higher light zones by Clade LLIV.",
"Potential clade occupancies for the southern temperate zone mirror the northern temperate zone occupancies.",

>>>>

#  This is Fig10.tiff
NichesMap_cap <- glue("**Potential Future Niches for *Prochlorococcus* clades in a warming ocean.**",
                      "The maps shows current latitudinal distribution limits for *Prochlorococcus* (between green dashed lines), along with tropical and temperate  latitudinal regions (yellow bands).",
"Light levels, photoperiods and spectral bands, potentially permissive for growth of *Prochlorococcus* clades are determined by the interactions of latitude, season and depth attenuation of light.",
"**A.** Potential clade occupancies under hypothetical Full Oxygen Zones (250 µM O~2~) are indicated for 10-50 m and 100 m depths, in temperate vs. tropical regions, under depth-resolved photoperiods, approximating Winter Solstice, Equinox and Summer Solstice. Clade HLI is excluded from temperate depths by short photoperiods at Winter Solstice. Clades LLII/III and LLIV are excluded from shallow depths by excessive light.",
"**B.** Oxygen Minimum Zones (25 µM O~2~, not shown on map) would drastically alter these potential occupancy patterns by excluding clade HLI from the temperate winter solstice at all depths, and by expanding the exploitation of higher light zones by Clade LLIV.",
"Potential clade occupancies for the southern temperate zone mirror the northern temperate zone occupancies.",
